---
title: "2026-02-26"
format: html
---

## Confidence Intervals

The SE can be used to derive CI (another way of describing a statistic's sampling distribution and it plays a central role in basic inferential statistics)

An interval around our estimate of mean of the sampling distribution for a particular statistic (typically a mean), and it gives us a range of values into which subsequent estimates of a statistic would be expected to fall some critical proportion of the time, if the sampling exercise were to be repeated

### Calculating Confidence Intervals

Theoretical confidence intervals (which are based on distributions with known mathematical properties) represent the range of values corresponding 

the value of the statistic +- critical balue * the standard error of the statistic

mean +- critical value * SE

CI = point estimare +- critical value x SE (critical value = how many standard deviations we're going based on % confidence??? So a 95%  confidence interval looks at 1.959964 standard deviations or SE's away from the mean??? (calculated via qnorm))

What is the theoretical 95% CI around the estimare of the mean of the following vector:

``` {r}
x <- c(2.9, 4.8, 8.9, -3.2, 9.1, -2.5, -0.9, -0.1, 2.8, -1.7)
m <- mean(x)
se <- sd(x)/sqrt(length(x))
ci <- m +qnorm(c(0.025, 0.975)) * se # or we can do it this way:
ci <- m + c(-1, 1) * qnorm(0.975) * se
```

## CI's for small sample sizes...

For n<30, use things like student's t distribution

Yher choice of the appropriate t distribution to use in any particular statistical test is based on the number of degrees of freedom (df), i.e., the number of individual components in the calculation of a 

``` {r}
n_boot <-  10000
boot <- vector()
n <- length(x)
for(i in 1:n_boot) {
  boot[[i]] <- 
}
boot <- do(n_boot) * 
```

## Estimating confidence intervals (redux)...

### If we know complete population...

Based on population mean, population SD, and sample size (to calculate expected SE) using qnorm()

``` {r}
ci <- popmean + qnorm(c(0.025, 0.975) * popSD/sqrt(n))
```

### If we can take multiple samples from the population...

Based on SD of sampling distribution (to estimate SE) using qnorm()

``` {r}
ci <- samplemean + qnorm(c(0.025, 0.975) * sd(samplingdistribution))
```

Directly from sampling distribution using quantile()

``` {r}

```

# Class 12: Hypothesis Testing

## 33 Null Hypothesis Significance Testing

Classical hypothesis testing typically involves formally stating a claim - the null hypothesis - which is then folllowed up by statistical evaluation of the null versus and alternative hypothesis.

The null hypothesis is interpreted as a baseline hypothesis and is the claim that is presumed to be true. That claim is typically that a particular value of a population parameter estimated by a sample statistic we have calculated is consistent with a particular null expectation. The alternative hypothesis is the conjecture that we are testing, usually that the sample statistic is inconsistent with a null expectation.

Typically, our null and alternative hypotheses are expressed something like as follows:

  
## To implement a hypothesis test, we need to...

Calculate a test statistic based on our data.

Calculate the p value assoc

In traditional parametric statistics, we make the assumption that the sampling distribution of our statistic of interest (e.g. a mean) talkes the form of a particular

## Process:

Specify the sample statistic we want to evaluate (e.g. the mean).

Specify the test statistic of interest and the form of the dsampling distriution for htat statistic (e.g. Z or T ans standard norma or T distribution for aprametric statistical tests).

T = (mean - se)/(se)

Calculate the tail probabiltiy 

``` {r}
f <- "https://raw.githubusercontent.com/difiore/ada-datasets/refs/heads/main/woolly-weights.csv"
d <- read.csv(f)
head(d)
m <- mean(d$weight)
sd <- sd(d$weight)
se <- sd/sqrt(length(d$weight))
t <- (m - 7.2)/se
ci <- m + c(-1, 1) * qnorm(0.975) * se # our sample mean is not within the 95% CI of the population; this means the p value calculated from the t statistic is likely to be <0.05
```



``` {r}

```



``` {r}

```



``` {r}

```