---
title: "2026-02-24"
format: html
---

## What is a sampling distribution?

### Sampling distribution of a statistic

### Estimating SE from a single sample

We can estimate the SE associated with samples of size n from a single sample of size n as the standard deviation of the sample divided by the square root of the sample size...

``` {r}
x <- rnorm(n = 10, mean = 10, sd = 2)
se <- sd(x)/sqrt(length(x))
print(x)
print(se)
```

### What happens to the spread of the sampling distribution, and thus to the SE, as the size of our samples increases? (e.g., change n = 10 to n = 1000)?

``` {r}
library(mosaic)

reps <- 1000
samp_dist_mean <- do(reps) * mean(rnorm(n = 1000, mean = 10, sd = 2))
histogram(~mean, data = samp_dist_mean)
se_mean <- sd(samp_dist_mean$mean)
```

``` {r}
reps <- 1000
samp_dist_median <- do(reps) * median(rnorm(n = 1000, mean = 10, sd = 2))
histogram(~median, data = samp_dist_median)
se_median <- sd(samp_dist_median$median)
```

``` {r}
reps <- 1000
samp_dist_sd <- do(reps) * sd(rnorm(n = 1000, mean = 10, sd = 2))
histogram(~sd, data = samp_dist_sd)
se_sd <- sd(samp_dist_sd$sd)
```

## Estimating standard error in a sample... 3 ways

### If true population variance is known (rare)...

(standard deviation of population)/sqrt(sample size)

### If we can sample population repeatedly to generate a sampling distribution...

Standard deviation of sampling distrpbution

### If we 

``` {r}
s <- rnorm(10, mean = 0, sd = 1)
sd(s)/sqrt(length(s))
```

# Class 11: Playing with Distributions

``` {r}
r <- rnorm(100, mean = 2, sd = 4)
cat("mean: ", mean(r))
cat("standard deviation: ", sd(r))
cat("standard error: ", sd(r)/sqrt(length(r)))
cat("expected standard error: ", 4/sqrt(100))
```

## Now repeat this 1000 etc times to get a sampling distribution

``` {r}
reps <- 1000
sdm <- do(reps) * mean(rnorm(n = 100, mean = 2, sd = 4))
cat("sampling distribution mean: ", mean(sdm$mean), "\n")
cat("standard error: ", sd(sdm$mean), "\n")
```

## Repeat this using a t distribution

``` {r}
plotDist("t", df = 100, xlab = "x", ylab = "Frequency", col = "red")
```

``` {r}
plotDist("t", df = 50, xlab = "x", ylab = "Frequency", col = "red")
```

``` {r}
plotDist("t", df = 25, xlab = "x", ylab = "Frequency", col = "red")
```

``` {r}
plotDist("t", df = 12, xlab = "x", ylab = "Frequency", col = "red")
```

``` {r}
plotDist("t", df = 6, xlab = "x", ylab = "Frequency", col = "red")
```

``` {r}
plotDist("t", df = 3, xlab = "x", ylab = "Frequency", col = "red")
```

``` {r}
plotDist("t", df = 1, xlab = "x", ylab = "Frequency", col = "red")
```

``` {r}
plotDist("t", df = 100, xlab = "x", ylab = "Frequency", col = "red")
plotDist("t", df = 50, xlab = "x", ylab = "Frequency", col = "red", add = TRUE)
plotDist("t", df = 25, xlab = "x", ylab = "Frequency", col = "red", add = TRUE)
plotDist("t", df = 12, xlab = "x", ylab = "Frequency", col = "red", add = TRUE)
plotDist("t", df = 6, xlab = "x", ylab = "Frequency", col = "red", add = TRUE)
plotDist("t", df = 3, xlab = "x", ylab = "Frequency", col = "red", add = TRUE)
plotDist("t", df = 1, xlab = "x", ylab = "Frequency", col = "red", add = TRUE)
```

``` {r}
rnorm(n = 100, mean = 2, sd = 4, ncp = 2)
```

## Repeat this exercise again using a beta distribution

``` {r}
plotDist("beta", shape1 = 2, shape2 = 4, xlab = "x", xlim = c(0,1), ylab = "Frequency", col = "red")
```

``` {r}
plotDist("beta", shape1 = 1, shape2 = 1, xlab = "x", xlim = c(0,1), ylab = "Frequency", col = "red")
```

``` {r}
plotDist("beta", shape1 = 0.5, shape2 = 2, xlab = "x", xlim = c(0,1), ylab = "Frequency", col = "red")
```

``` {r}
plotDist("beta", shape1 = 2, shape2 = 2, xlab = "x", xlim = c(0,1), ylab = "Frequency", col = "red")
```

``` {r}
plotDist("beta", shape1 = 4, shape2 = 2, xlab = "x", xlim = c(0,1), ylab = "Frequency", col = "red")
```

``` {r}
plotDist("beta", shape1 = 0.5, shape2 = 0.5, xlab = "x", xlim = c(0,1), ylab = "Frequency", col = "red")
```

``` {r}
plotDist("beta", shape1 = 2, shape2 = 0.5, xlab = "x", xlim = c(0,1), ylab = "Frequency", col = "red")
```

``` {r}
reps <- 1000
b <- do(reps) * mean(rbeta(n = 100, shape1 = 0.3, shape2  = 4))
cat("mean: ", mean(b$mean), "\n")
cat("standard deviation: ", sd(b$mean), "\n")
cat("standard error: ", sd(b$mean)/sqrt(length(b$mean)), "\n")
cat("expected SE: ")
histogram(~mean, data = b)
```

## Key Functions for Distributions

r_(n = ) draws random samples of size n from a given distribution

p_(q = ) returns the quantile associated with a given value of X; i.e., returns the value of the cumulative density function up to that value of

q_(p = ) returns the value of X at a given quantile through the distribution; i.e., returns the value of the inverse cumulative density dunction (cdf)

d_(x = ) returns the value of the probability density function (pdf) at the values of x

What values of X are associated with the 0.025 and 0.975 quantiles of a standard normal distribution (which woudl be the boundaries of the central 95% of a given distribution)?

``` {r}
qnorm(p = c(0.025, 0.975), mean = 0, sd = 1)
```

What values of X are associated with the .0025 and 0.975 quantiles of a beta distribution with shape1 = 2 and shape2 = 4?

``` {r}
qbeta(p = c(0.025, 0.975), shape1 = 2, shape2 = 4)
```

``` {r}
lower <- qbeta(0.025, shape1 = 2, shape2 = 4)
upper <- qbeta(0.975, shape1 = 2, shape2 = 4)
curve(dbeta(x, shape1 = 2, shape2 = 4), from = 0, to = 1, lwd = 2)
x <- seq(lower, upper, length = 500)
polygon(c(lower, x, upper), c(0, dbeta(x, shape1 = 2, shape2 = 4), 0), col = "lightblue")
abline(v = c(lower, upper), col = "red", lwd = 2)
```

``` {r}
pbeta(q = 0.2, shape1 = 2, shape2 = 4)
```

``` {r}
qbeta(p = 0.5, shape1 = 2, shape2 = 4)
```