---
title: "2026-03-05"
format: html
---

## Permutation Distributions

``` {r}
library(tidyverse)
library(dplyr)
library(mosaic)
library(infer)
d <- tibble(read.csv("https://raw.githubusercontent.com/difiore/ada-datasets/refs/heads/main/tbs-2006-2008-ranges.csv"))

```

``` {r}
n_perm <- 10000
perm_diff <- vector()
for(i in 1:n_perm) {
  d %>%
    mutate(sex = sample(sex)) %>%
    summarize(diff = mean(kernel95[sex == "M"]) - mean(kernel95[sex == "F"])) %>%
    pull(diff) -> perm_diff[[i]]
}
histogram(perm_diff)
```

permuted difference between means

``` {r}
perm_diff <- {
  do(n_perm) * d %>%
    mutate(sex = sample(sex)) %>%
    summarize(diff = mean(kernel95[sex == "M"]) - mean(kernel95[sex == "F"]))
} %>%
  pull(diff)
histogram(perm_diff)
```

``` {r}
(p_val <- sum(abs(perm_diff) >= abs(obs_diff))/length(perm_diff)) # how many times is our permuted value bigger than our observed value?
# or
(p_val <- mean(abs(perm_diff) >= abs(obs_diff)))
```

``` {r}
d <- d %>%
  specify(formula = kernel95 ~ sex)
d <- d %>%
  hypothesize(null = "independence")
perm <- d %>%
  generate(reps = 1000, type = "permute")
perm_diff <- perm %>%
  calculate(stat = "diff in means", order = c("M", "F"))
obs_diff <- d %>%
  summarize(diff = mean(kernel95[sex == "M"]) - mean(kernel95[sex == "F"])) %>%
  pull(diff)
obs_diff <- d %>%
  specify(kernel95 ~ sex) %>%
  calculate(stat = "diff in means", order = c("M", "F"))
visualize(perm_diff, bins = 20) +
  shade_p_value(obs_stat = obs_diff, direction = "both")
```

``` {r}
get_p_value(perm_diff, obs_diff, direction = "both")
```

# Class 13: Introduction to Regression

``` {r}
library(lmodel2)
library(sjPlot)
library(broom)
library(tidyverse)
library(manipulate)
library(patchwork)
library(infer)
```

## Regression -- A common form of data modeling

The basic premise of modeling is to explore the relationship between the outcome variable (y) and one or more explanatory variables (x)

### Simple (general) linear regression

Outcome is a continuous numerical variable, single predictor that is either numerical or categorical

### Multiple (general) linear regression

Outcome is a continuous numerical variable, multiple predictors that are either numerical or categorical

### ANOVA/ANCOVA

Focuses on categorical predictors

ANOVA - one or more categorical independent variables

ANCOVA - a combination or categorical and continuous independent variables

### "Generalized" linear regression

Allows for binary, categorical, count variables as outcomes

## Before modeling, start with exploratory data analysis

Univariate summary statistics, e.g. skinm() from {skimr}

Bivariate summary statistics

- Covariance expresses how much two numeric variables "change together" and whether that change is positive or negative

- Correlation coefficient is a standardized form of the covariance that summarizes, on a scale from -1 to +1, both the strength and direction of a relationship

- Covariance between two variables measured on a sample is the product of the deviations of each of those variables from their respective means divided by (sample size - 1)

cov(x, y) = sum((x - mean(x))(y - mean(y))/(n - 1))

- Correlation Coefficient two is a standardized covariance (divide by the product of the standard deviations of the two variables)

cor(x, y) = cov(x, y)/(sd(x) * sd(y))

### Programming exercise...

``` {r}
d <- read_csv("https://raw.githubusercontent.com/difiore/ada-datasets/refs/heads/main/zombies.csv", col_names = TRUE)
head(d)
```
``` {r}
z_plot <- ggplot(data = d, aes(x = weight, y = height)) +
  geom_point()
z_plot
```

``` {r}
z_cov <- cov(d$weight, d$height)
z_cov
```

``` {r}
z_cor <- cor(d$weight, d$height)
z_cor
```

``` {r}
z_cov_hand <- sum(((d$height - mean(d$height)(d$weight - mean(d$weight)))/(length(d$height) - 1)))
z_cov_hand
```

## Main Purposes for Regression

To use one or more variables to predict the value of an outcome variable y

To explicitly describe and quantify the relationship between the outcome variable y and a set of explanatory variables x, determine the significance of any relationships, generate measures summarizing these relationships (e.g. regression coefficients), and (possibly) identify causal relatinoships between the variables

To develop and choose among different models of the relationships between the variables of interest

To do analyses of covariation among sets of variables to identify/explore their relative explanatory power

## Formula notation for regression

We typically model the outcome variable y "as a linear function" of the explanatory/predictor variables

y(i) = beta(0) + beta(1)chi(i) + epsilon(i)

The beta values in this equation arerreferred to as "regression coefficients", and it is those coefficients that our analysis is trying to estimate, while minimizing, according to some criterion, the error term

y(hat) = beta(1)x + beta(0)

## One common criterion is ORDINARY LEAST SQUARES

We want to find coefficients that minimize "the residuals", i.e., the difference between observed and expected y values

sum(y - y(hat)^2)