---
title: "2026-03-12"
format: html
---

## Maximum Likelihood Estimation

Probability of seeing a value (mu) according to its mean (x) and standard deviation (sigma^2)

AKA a probability density function

P(x|mu, sigma) = (1/sqrt(2 * pi * sigma^2)) * e^(-(x - mu)^2/(2 * sigma^2))

``` {r}
library(bbmle)
library(maxLik)
```

``` {r}
val <- c(41, 70, 10)
mean <- 50
sd <- 10
l <- dnorm(val, mean = mean, sd = sd) # vector of likelihoods of each value
l
```

``` {r}
l[1] * l[2] * l[3] # product of likelihoods
```

``` {r}
ll <- log(l) # can be easier to work with
ll[1] + ll[2] + ll[3] # product of likelihoods = sum of log likelihoods
```

``` {r}
set.seed(0)
d <- rnorm(100, 50, 10)
head(d)
```

``` {r}
l <- dnorm(d, mean, sd)
ll <- log(l)
ll <- sum(ll)
l <- exp(ll)
l
```

``` {r}
mean <- 65
sd <- 20
l <- dnorm(d, mean, sd)
ll <- log(l)
ll <- sum(ll)
l <- exp(ll)
l
```

``` {r}
minuslog1 <- function(mu, sigma, verbose = TRUE) { # automatically produces log likelihood as a positive number
  ll = sum(dnorm(d, mean = mu, sd = sigma, log = TRUE))
  nll <- -1 * ll
  return(nll)
}
m <- mle2( # maximum likelihood estimate 2
  minuslogl = minuslog1,
  start = list(mu = 0, sigma = 1),
  method = "SANN" # simulated annealing method of optimization, one of several options
)
summary(m)
```

``` {r}
library(tidyverse)
d <- read_csv("https://raw.githubusercontent.com/difiore/ada-datasets/main/zombies.csv", col_names = TRUE)
m <- lm(data = d, height ~ weight) # makes  a linear model of the data (line of best fit, mean, etc)
SSY <- sum((m$model$height - mean(m$model$height))^2) # height - mean(height); number of degrees of freedom for the total sum of squares (SSY) is n - 1
SSR <- sum((m$fitted.values - mean(m$model$height))^2) # degrees of freedom for regression sum of squares (SSR) is equal to the number of predictor variables (p)
SSE <- sum((m$model$height - m$fitted.values)^2) # df for the error sum of squares (SSE) is equal to n - (p + 1) (AKA n - p - 1)
df_y <- nrow(d) - 1
df_regression <- p
df_error <- nrow(d) - p - 1

MSY <- SSY/df_y
MSR <- SSR/df_regression
MSE <- SSE/df_error
```

## ANOVA tables: F ratio

fratio = MSR/MSE

ratio of the variance explained by the regression model to the remaining unexplained variance

We can then test the overall significance of our regression model by evaluating our F ratio test statistic against an F distribution

``` {r}
for(i in 1:3) {
  cat("meow ")
  for(j in 1:3) {
    cat("Meow ")
    for(k in 1:3) {
      cat("mEow ")
      for(l in 1:3) {
        cat("meOw ")
        for(m in 1:3) {
          cat("meoW ")
          for(n in 1:3) {
            cat("MEow ")
            for(o in 1:3) {
              cat("MeOw ")
              for(p in 1:3) {
                cat("MeoW ")
                for(q in 1:3) {
                  cat("mEOw ")
                  for(r in 1:3) {
                    cat("mEoW ")
                    for(s in 1:3) {
                      cat("meOW ")
                      for(t in 1:3) {
                        cat("MEOw ")
                        for(u in 1:3) {
                          cat("MEoW ")
                          for(v in 1:3) {
                            cat("MeOW ")
                            for(w in 1:3) {
                              cat("mEOW ")
                              for(x in 1:3) {
                                cat("MEOW ")
                              }
                            }
                          }
                        }
                      }
                    }
                  }
                }
              }
            }
          }
        }
      }
    }
  }
}
```