---
title: "2026-03-03"
format: html
---

## One Sample T test

T = (mean - expectation)/(standard error)

``` {r}
ci <- m + qt(p = c(0.025, 0.975), ncp = 0, df = n - 1) * se
```

Does CI include the expected mean? NO

What is the two-tailed probability of getting a statistic as far away from the expected value (0) as it is by chance?

Because this is a two-tailed test, we need to calculated the probability of seeing a T statistic as far from the mean in either the positive or negative direction. So, we sum the probabilities associated with the upper and lower tails...

``` {r}
library(mosaic)
f <- "https://raw.githubusercontent.com/difiore/ada-datasets/refs/heads/main/woolly-weights.csv"
d <- read.csv(f)
```

``` {r}
m <- mean(d$weight)
se <- sd/sqrt(length(d$weight))
mu <- 7.2
t_stat <- (m-mu)/se
n <- length(d$weight)
```

``` {r}
ci <- m + qt(p = c(0.025, 0.975), ncp = 0, df = n - 1) * se
p_lower <- pt(-1 * abs(t_stat), df = n - 1)
p_upper <- pt(1 * abs(t_stat), df = n-1)
p <- c(p_lower, p_upper)
```

## Programming Exercise: A Two-Sample T test

``` {r}
d <- read.csv("https://raw.githubusercontent.com/difiore/ada-datasets/refs/heads/main/tbs-2006-2008-ranges.csv")
s <- d %>%
  select(id, sex, kernel95) %>%
  group_by(sex) %>%
  summarize(avg = mean(kernel95),
            sd = sd(kernel95),
            se = sd/sqrt(nrow(d))
            )
s
```

``` {r}
p <- ggplot(data = d, aes(x = sex, y = kernel95)) +
  geom_boxplot() +
  geom_point() +
  geom_jitter()
p
```

``` {r}
males <- d %>% filter(sex == "M")
females <- d %>% filter(sex == "F")

n_boot <- 10000
boot_males <- (do(n_boot) *
                 mean(sample(
                   males$kernel95,
                   nrow(males),
                   replace = TRUE))) %>%
  pull(mean)
histogram(boot_males)
plotDist("norm", mean(boot_males), sd(boot_males), add = TRUE) # superimposes a normal distribution curve to see how "normal" your distribution is
```

``` {r}
boot_females <- (do(n_boot) *
                 mean(sample(
                   females$kernel95,
                   nrow(females),
                   replace = TRUE))) %>%
  pull(mean)
histogram(boot_females)
plotDist("norm", mean(boot_females), sd(boot_females), add = TRUE)
```

``` {r}
m_males <- mean(boot_males)
sd_males <- sd(boot_males)

ci_boot_m <- quantile(boot_males, probs = c(0.025, 0.975))
ci_norm_m <- m_males + qnorm(c(0.025, 0.975)) * sd_males
```

## The write code to do the following inferential analysis...

Is there a significant difference in mean home range size of males and females?

What is H(o)? What is H(a)?

T = (mean of group 2 - mean of group 1 - expected difference)/sqrt((group 1 standard deviation)^2/n1 + (group 2 standard deviation)^2/n2)

``` {r}
t_num <- mean(males$kernel95) - mean(females$kernel95)
t_denom <- sqrt(var(males$kernel95)/nrow(males) + var(females$kernel95)/nrow(females))
t <- t_num/t_denom
```

``` {r}
df_num <- (var(males$kernel95)/nrow(males) + var(females$kernel95)/nrow(females))^2
df_denom <- ((var(males$kernel95)/nrow(males))^2)/(nrow(males)-1) + ((var(females$kernel95)/nrow(females))^2)/(nrow(females)-1)
df <- df_num/df_denom
```